---
title: "microbiomeDataSets"
output: 
  BiocStyle::html_document:
    fig_height: 7
    fig_width: 10
    toc: yes
    toc_depth: 2
    number_sections: true
vignette: >
  %\VignetteIndexEntry{microbiomeDataSets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(microbiomeDataSets)
```

# Microbiome example data sets

The data sets are primarily named by the first author of the
associated publication, together with a descriptive suffix. Aliases
are provided for some of the data sets.

A table of the available data sets is available through the `availableDataSets`
function.

```{r}
availableDataSets()
```

The data is usually loaded as a `r Biocpkg("TreeSummarizedExperiment")` or a
`r Biocpkg("MultiAssayExperiment")`, if microbiome associated data was used
in the respective study. For more information on how to use these objects,
please refer to the vignettes of the packages.

## HIT Chip Data

The following data sets are based on the Human Intestinal Tract (HIT)Chip
phylogenetic 16S microarray [(Rajilić-Stojanović _et al._
2009)](https://doi.org/10.1111/j.1462-2920.2009.01900.x). This profiling
technology differs from the more widely used 16S rRNA amplicon sequencing.

Since the data is also available in phyloseq format through the microbiome R
package and are referenced usually via aliases, they are described in more
detail.

### Intestinal microbiota profiling of 1006 Western adults

This data set from [Lahti et al. Nat. Comm. 5:4344, 2014](http://www.nature.com/ncomms/2014/140708/ncomms5344/full/ncomms5344.html)
comes with 130 genus-like taxonomic groups across 1006 western adults with no
reported health complications. Some subjects have also short time series.

```{r atlasdata, warning=FALSE, message=FALSE}
LahtiWAData()

# Alias
# atlas1006()
```

### Diet swap between Rural and Western populations

A two-week diet swap study between western (USA) and traditional (rural Africa)
diets, reported in [O'Keefe et al. Nat. Comm. 6:6342, 2015](http://dx.doi.org/10.1038/ncomms7342).

```{r dietswap}
OKeefeDSData()

# Alias
# dietswap()
```

### Intestinal microbiota versus blood metabolites

Data set from [Lahti et al. PeerJ 1:e32, 2013](https://doi.org/10.7717/peerj.32)
characterizes associations between human intestinal microbiota and blood serum
lipids. Note that this data set contains an additional assay of lipid species,
and is therefore provided as `MultiAssayExperiment` object.

```{r peerj}
LahtiMLData()

# Alias
# peerj32()
```
